Models of Lambek calculus with subexponentials

### 2019-07-05 PaperDraft.tex
fixes

  $\mathcal{F}_{\Sigma} ::= Tp \: | \: (\mathcal{F}_{\Sigma} / \mathcal{F}_{\Sigma}) \: | \: (\mathcal{F}_{\Sigma} \backslash
  \mathcal{F}_{\Sigma}) \: | \: (\mathcal{F}_{\Sigma} \bullet \mathcal{F}_{\Sigma}) \: | \: (\mathcal{F}_{\Sigma} \lor \mathcal{F}_{\Sigma}) \: | (\mathcal{F}_{\Sigma} \land \mathcal{F}_{\Sigma}) \: | ({!_s \mathcal{F}_{\Sigma}})_{s \in
  \Sigma} $
\end{defin}

Let us introduce inference polymodal inference rules for non-commutative subexponentials. One can apply substructural
rules only if there is the relevant index on the current modality. Also, the right introduction rule is a sort of the
generalised right $\Box$-introduction a la $K4$. Modality $!_s$ might be introduced on the right only if its index is less than any other subexponential index from the antecedent. That is, if we have already used stronger modality, then we may
apply the weaker one.

@@

Let us consider the current proof-theoretical and algorithmic results on Lambek calculus with additives and subexponentials.
The first of all, the cut rule is admissible. Generally, this calculus is undecidable, but the fragment without non-local
contraction is PSPACE-complete. These results were obtained by Kanovich, Kuznetsov, Nigam, and Scedrov \cite{Sc}.

@@

\begin{itemize}
\item Let $A$ be a semigroup (monoid), then $\langle \mathcal{P}(A), \cdot, \subseteq \rangle$
is a free (unital) quantale.
\item Let $R$ be a ring and $Sub(R)$ be a set of additive subgroups of $R$.
We define $A \cdot B$ as an additive subgroup generated by finite sums of products $ab$ and order is defined by inclusion.
\item Any locale is a quantale with $\cdot = \wedge$.
\end{itemize}

@@

Then, if $f : \mathcal{Q}_1 \to \mathcal{Q}_2$ is a quantale homomorphism, then $f(\mathcal{S}) \subseteq \mathcal{Q}_2$ is a subquantale of $\mathcal{Q}_2$.
\end{prop}

@@

  Follows from the same statements for semigroup and semilattice homomorphisms.
\end{proof}

@@

  \item A quantic conucleus is called unital, if $\forall a \in Q, \Box a \leq e$.
\end{enumerate}
\end{defin}

@@

  \begin{enumerate}
    \item Let $S \subseteq \mathcal{Q}$ be a subquantale, then $I : \mathcal{Q} \to \mathcal{Q}$, such that $\Box a = \bigvee \{ s \in S \: | \: s \leq a \}$, is an open modality.
    \item Let $S_1$ and $S_2$ are subquantales such that $\mathcal{S}_1 \subseteq \mathcal{S}_2 \subseteq \mathcal{Q}$. Then $I_{\mathcal{S}_1} (a) \leq I_{\mathcal{S}_1} (a)$, where $I_{\mathcal{S}_i}, i = 1,2$ are quantic conucleis defined according to the previous part.
  \end{enumerate}
\end{lemma}

@@

\begin{lemma} Let $I$ be a quantic conucleus, then for all $a \in \mathcal{Q}$, $I a = \bigvee \{ q \in Q_{I} \: | \: s \leq a \}$, where $Q_{I} = \{ a \in \mathcal{Q} \: | \: I a = a \}$.
\end{lemma}

@@

So far as $Q_{I} = \{ a \in \mathcal{Q} \: | \: I a = a \}$ is a subquantale, then $I' a = \bigvee \{ q \in Q_{I} \: | \: s \leq a \}$ is a quantic conucleus. Let us show that these operations are equavalent.

@@

$\begin{array}{lll}
&\bigvee \{ q \in Q_{I} \: | \: s \leq a \} = & \\
& \:\:\:\:\:\:\:\: \text{$s \in Q_{I}$, so s is a fixpoint of $I$} & \\
&\bigvee \{ q \in Q_{I} \: | \: I s \leq a \} = & \\
& \:\:\:\:\:\:\:\: \text{By equivalence}& \\
&\bigvee \{ q \in Q_{I} \: | \: I s \leq I a \} = & \\
& \:\:\:\:\:\:\:\: \text{By fixpoint} & \\
&\bigvee \{ q \in Q_{I} \: | \: s \leq I a \} \leq I a& \\
\end{array}$

@@

\begin{enumerate}
  \item $I a = \bigvee \{ s \in S \: | \: s \leq a, s \in \mathcal{Z}(\mathcal{Q}) \}$ is a central quantic conucleus
  \item $I a = \bigvee \{ s \in S \: | \: s \leq a, s \leq \mathds{1} \}$ is an unital conucleus
  \item $I a = \bigvee \{ s \in S \: | \: s \leq a, \forall b \in Q, b \cdot s \vee s \cdot b \leq s \cdot b \cdot s\}$ is a generalised square-increasing quantic conucleus
  \item Operations with combinations of conditions above
\end{enumerate}
\end{prop}

@@

  $\sigma(s_i) = \begin{cases}
  I_i : \mathcal{Q} \to \mathcal{Q} \text{ s.t. } \forall a \in Q, I_i(a) = \{ s \in S_i \: | \: s \leq a\},
  \\ \:\:\:\: \text{if $s_i \notin W \cap C \cap E$} \\
  I_i : \mathcal{Q} \to \mathcal{Q} \text{ s.t. } \forall a \in Q, I_i(a) = \{ s \in S_i \: | \: s \leq a, s \leq \mathds{1}\},
  \\ \:\:\:\: \text{if $s_i \in W$} \\
  I_i : \mathcal{Q} \to \mathcal{Q} \text{ s.t. } \forall a \in Q, I_i(a) = \{ s \in S_i \: | \: s \leq a, s \in \mathcal{Z}(\mathcal{Q}) \},
  \\ \:\:\:\: \text{if $s_i \in E$} \\
  I_i : \mathcal{Q} \to \mathcal{Q} \text{ s.t. } \forall a \in Q, I_i(a) = \{ s \in S_i \: | \: s \leq a, \forall b, b \cdot s \vee s \cdot b \leq s \cdot b \cdot s \},
  \\ \:\:\:\: \text{if $s_i \in E$} \\
  \text{otherwise, if $s_i$ belongs to some intersection of subsets, then we combine the relevant conditions } \\
  \end{cases}$
\end{defin}

\begin{defin} Let $\mathcal{Q}$ be an unital quantale, $f : Tp \to \mathcal{Q}$ a valuation and $\sigma : I \to CN(\mathcal{Q})$ a subexponential interpretation, then interpretation is defined inductively:

@@

\begin{prop} $s \in A = \{ \mathcal{W}, \mathcal{C}, \mathcal{E} \}$ iff $\sigma(s)$ is a unital (central or square increasing) modality.
\end{prop}

@@

  Let $\Sigma$ be a subexponential signature and $\sigma : \Sigma \to CN(\mathcal{Q})$ for a contravariant map
  $S : \Sigma \to Sub(\mathcal{Q})$, then:
  \begin{enumerate}
  \item For all $s_1, s_2 \in \Sigma$, if $s_1 \preceq s_2$, then $\sigma(s_2) \leq \sigma(s_1)$.
  In other words, $\sigma$ is a contravariant functor from $\Sigma$ to $Sub(\mathcal{Q})$.
  \item Let $A = \{ \mathcal{W}, \mathcal{C}, \mathcal{E} \}$, then $A$ is downwardly closed
  $\sigma(A) = \{ \sigma(s) \: | \: s \in A \}$ is downwardly closed, where $A = \{ \mathcal{W}, \mathcal{C}, \mathcal{E} \}$.
  \item If $s \in \mathcal{W} \cap \mathcal{C}$, then $\sigma(s)$ is a central quantic conucleus.
\end{enumerate}
\end{lemma}

>>>>

  $\mathcal{F}_{\Sigma} ::= Tp \: | \: {\bf 1} \: | \: (\mathcal{F}_{\Sigma} / \mathcal{F}_{\Sigma}) \: | \: (\mathcal{F}_{\Sigma} \backslash
  \mathcal{F}_{\Sigma}) \: | \: (\mathcal{F}_{\Sigma} \bullet \mathcal{F}_{\Sigma}) \: | \: (\mathcal{F}_{\Sigma} \lor \mathcal{F}_{\Sigma}) \: | (\mathcal{F}_{\Sigma} \land \mathcal{F}_{\Sigma}) \: |
   \: ({!_s \mathcal{F}_{\Sigma}})_{s \in \Sigma} $
\end{defin}

Let us introduce the polymodal inference rules for non-commutative subexponentials. One can apply substructural
rules only if there is the relevant index on the current modality. Also, the right introduction rule is a sort of the
generalised right $\Box$-introduction a la the modal logic ${\bf K}4$. Modality $!_s$ might be introduced on the right only if its index is less than any other
subexponential index from the antecedent. That is, if we have already used stronger modality, then we may apply the weaker one.

@@

Let us consider the current proof-theoretical and algorithmic results on Lambek calculus with additives and subexponentials.
The first of all, the cut rule is admissible. Generally, this calculus is undecidable, but the fragment without non-local
contraction is PSPACE. These results were obtained by Kanovich, Kuznetsov, Nigam, and Scedrov \cite{Sc}.

@@

\begin{itemize}
\item Let $S$ be a semigroup (monoid), then $\langle \mathcal{P}(S), \cdot, \subseteq \rangle$
is a free (unital) quantale.
\item Let $\mathcal{R}$ be a ring and $Sub(\mathcal{R})$ be a set of additive subgroups of $R$.
We define $A \cdot B$ as an additive subgroup generated by finite sums of products $ab$ and order is defined by inclusion.
\item Any locale is a quantale with $\cdot = \wedge$.
\end{itemize}

@@

Then, if $f : \mathcal{Q}_1 \to \mathcal{Q}_2$ is a quantale homomorphism, then $f(\mathcal{S}) \subseteq \mathcal{Q}_2$ is a subquantale of $\mathcal{Q}_2$.

That is, a homomorphic image of subquatale is a subquantale.
\end{prop}

@@

  Obviously.
\end{proof}

@@

  \item A quantic conucleus is called unital, if $\forall a \in Q, I a \leq e$.
\end{enumerate}
\end{defin}

@@

  \begin{enumerate}
    \item Let $S \subseteq \mathcal{Q}$ be a subquantale, then the operation $I : \mathcal{Q} \to \mathcal{Q}$, such that $I a = \bigvee \{ q \in S \: | \: q \leq a \}$, is an open modality.
    \item Let $S_1$ and $S_2$ are subquantales such that $\mathcal{S}_1 \subseteq \mathcal{S}_2 \subseteq \mathcal{Q}$. Then $I_{\mathcal{S}_1} (a) \leq I_{\mathcal{S}_1} (a)$, where $I_{\mathcal{S}_i}, i = 1,2$ are quantic conucleis defined according to the previous part.
  \end{enumerate}
\end{lemma}

@@

\begin{lemma} Let $I$ be a quantic conucleus, then for all $a \in \mathcal{Q}$, $I a = \bigvee \{ q \in Q_{I} \: | \: q \leq a \}$, where $Q_{I} = \{ a \in \mathcal{Q} \: | \: I a = a \}$.
\end{lemma}

@@

So far as $Q_{I} = \{ a \in \mathcal{Q} \: | \: I a = a \}$ is a subquantale, then $I' a = \bigvee \{ q \in Q_{I} \: | \: q \leq a \}$ is a quantic conucleus. Let us show that these operations are equavalent.

@@

$\begin{array}{lll}
&\bigvee \{ q \in Q_{I} \: | \: q \leq a \} = & \\
& \:\:\:\:\:\:\:\: \text{$s \in Q_{I}$, so s is a fixpoint of $I$} & \\
&\bigvee \{ q \in Q_{I} \: | \: I q \leq a \} = & \\
& \:\:\:\:\:\:\:\: \text{By equivalence}& \\
&\bigvee \{ q \in Q_{I} \: | \: I q \leq I a \} = & \\
& \:\:\:\:\:\:\:\: \text{By fixpoint} & \\
&\bigvee \{ q \in Q_{I} \: | \: q \leq I a \} \leq I a& \\
\end{array}$

@@

\begin{enumerate}
  \item $I a = \bigvee \{ q \in S \: | \: q \leq a, q \in \mathcal{Z}(\mathcal{Q}) \}$ is a central quantic conucleus
  \item $I a = \bigvee \{ q \in S \: | \: q \leq a, q \leq \mathds{1} \}$ is an unital conucleus
  \item $I a = \bigvee \{ q \in S \: | \: q \leq a, \forall b \in Q, b \cdot s \vee s \cdot b \leq s \cdot b \cdot s\}$ is a generalised square-increasing quantic conucleus
  \item Operations with combinations of conditions above
\end{enumerate}
\end{prop}

@@

  $\sigma(s_i) = \begin{cases}
  I_i : \mathcal{Q} \to \mathcal{Q} \text{ s.t. } \forall a \in Q, I_i(a) = \{ q \in S_i \: | \: q \leq a\},
  \\ \:\:\:\: \text{if $s_i \notin W \cap C \cap E$} \\
  I_i : \mathcal{Q} \to \mathcal{Q} \text{ s.t. } \forall a \in Q, I_i(a) = \{ q \in S_i \: | \: q \leq a, q \leq \mathds{1}\},
  \\ \:\:\:\: \text{if $s_i \in W$} \\
  I_i : \mathcal{Q} \to \mathcal{Q} \text{ s.t. } \forall a \in Q, I_i(a) = \{ q \in S_i \: | \: q \leq a, s \in \mathcal{Z}(\mathcal{Q}) \},
  \\ \:\:\:\: \text{if $s_i \in E$} \\
  I_i : \mathcal{Q} \to \mathcal{Q} \text{ s.t. } \forall a \in Q, I_i(a) = \{ q \in S_i \: | \: q \leq a, \forall b, b \cdot q \vee q \cdot b \leq q \cdot b \cdot s \},
  \\ \:\:\:\: \text{if $s_i \in E$} \\
  \text{otherwise, if $s_i$ belongs to some intersection of subsets, then we combine the relevant conditions } \\
  \end{cases}$
\end{defin}

\begin{defin} Let $\mathcal{Q}$ be an unital quantale, $f : Tp \to \mathcal{Q}$ a valuation and $\sigma : \Sigma \to CN(\mathcal{Q})$ a subexponential interpretation, then interpretation is defined inductively:

@@

\begin{prop} $s \in A = \mathcal{W}, \mathcal{C}, \mathcal{E}$ iff $\sigma(s)$ is a unital (central or square increasing) modality.
\end{prop}

@@

  Let $\Sigma$ be a subexponential signature and $\sigma : \Sigma \to CN(\mathcal{Q})$ for a contravariant map
  $S : \Sigma \to Sub(\mathcal{Q})$, then:
  \begin{enumerate}
  \item For all $s_1, s_2 \in \Sigma$, if $s_1 \preceq s_2$, then $\sigma(s_2) \leq \sigma(s_1)$.
  In other words, $\sigma$ is a contravariant functor from $\Sigma$ to $Sub(\mathcal{Q})$.
  \item Let $A = \mathcal{W}, \mathcal{C}, \mathcal{E}$, then $A$ is downwardly closed
  $\sigma(A) = \{ \sigma(s) \: | \: s \in A \}$ is downwardly closed.
  \item If $s \in \mathcal{W} \cap \mathcal{C}$, then $\sigma(s)$ is a central quantic conucleus.
\end{enumerate}
\end{lemma}
